#First
##Packages used dada2, vegan, ggplot2 and plyr
```{r}
library(dada2)
library(ggplot2)
library(vegan)
library(plyr)
```

## created and set pathway to file locations adn reed in the files
```{r include=FALSE}
path <- "c:/Users/spencer/Desktop/Data_Course/Data_Course_mcgee/Final_Project/SpencerMcGee_Seqs_and_Metadata/"
list.files(path)

setwd(path)

snmeta = read.csv("Snail_MetaData.csv")

fq.files = dir(path = getwd(), full.names = TRUE, pattern = ".fastq")
plotQualityProfile(fq.files[12])
```

## get filenames for files without path and filter them and so you can place them in a new forlder
```{r warning=FALSE, include=FALSE}
filt.files = dir(path = getwd(), full.names = FALSE, pattern = ".fastq")

#dir.create(file.path(getwd(),"filtered"))
```

# stringent quality filter to remove N's etc, and place the filtered files of the fastq to it own folder and Dereplicate the fastq files then learn the error rates
```{r}
for(i in filt.files){
  fastqFilter(i,paste0(getwd(),"/filtered/",i,".filt"))
  
}
```
```{r}
setwd("C:/Users/spencer/Desktop/Data_Course/Data_Course_mcgee/Final_Project")
setwd(file.path(paste0(getwd(),"/filtered")))
filtered = dir(path = getwd(), pattern = ".filt")
dr = derepFastq(filtered)
err = learnErrors(dr)
```

##plot errors then clean the data nd create a Sequence table
```{r}
plotErrors(err)

clean = dada(dr,err)

SeqTable = makeSequenceTable(clean)
```


```{r eval=FALSE, include=FALSE}
# #taxonomy = assignTaxonomy(SeqTable, refFasta = "c:/Users/spencer/Desktop/Data_Course/Data_Course_mcgee/Final_Project/sh_general_release_dynamic_01.12.2017.fasta")
# 
# taxa.print <- taxonomy # Removing sequence rownames for display only
# rownames(taxa.print) <- NULL
# taxa.print
# 
# which(names(as.data.frame(SeqTable)) %in% row.names(as.data.frame(taxonomy)))
# RSV = as.data.frame(SeqTable)
# 
# taxa = as.data.frame(taxonomy)
# View(RSV)
# names(RSV) <- paste0(taxa$Family,"_",taxa$Genus)
# View(RSV)
```

##created a otu table from the Sequence table the from that create a data frame
```{r}
otu = SeqTable
t_otu = as.data.frame(otu)
```

##we then check the diversity of the table, see if there is anything that is identical then run adonis to see what is important.
```{r}
diversity(t_otu)

identical(row.names(t_otu), as.character(snmeta$InputFileName))

adonis(t_otu ~ snmeta$SiteName)
```
##we can then plot what we find.
```{r echo=FALSE}
#summary(colSums(t_otu))

summary(t_otu[,1])
plot(t_otu)

NMDS = metaMDS(t(t_otu))
NMDS$points
MDS1 = NMDS$points[,1]
MDS2 = NMDS$points[,2]

plot(MDS1,MDS2)
```

##we can even plot it in ggplot2 and creat a heat map.
```{r include=FALSE}
df = data.frame(MDS1 = MDS1, MDS2=MDS2, sitename = snmeta$SiteName)

ggplot(df, aes(x=MDS1, y=MDS2,color=sitename)) +
  geom_point() +
  stat_ellipse()
```
```{r include=FALSE}
Heatmap = heatmap(as.matrix(t_otu))
```
```{r}
Heatmap_t = heatmap(as.matrix(t(t_otu)))
```

##we can then color it and apply it to the heat map.
```{r include=FALSE}
color = as.character(mapvalues(snmeta$SiteName, from = c("D_Sievert","Ekahanui","Kaala_Bog","Kaaikukai","Skeets_Pass"), 
                               to = c("Blue","Red","Green","Yellow","Black")))

heatmap2 = heatmap(as.matrix(t(t_otu)), ColSideColors = color, col = gray.colors(100)) 
```
```{r echo=FALSE}
png(filename = "c:/Users/spencer/Desktop/Data_Course/Data_Course_mcgee/Final_Project/Heatmap.png")
heatmap(as.matrix(t(t_otu)), ColSideColors = color, col = gray.colors(100))
dev.off()
```